---
title: "Estimating synchronous and compensatory dynamics in grassland community using a novel “copula” approach"
author: "Shyamolina Ghosh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
#date: "October 12, 2018"
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=1.5cm"

output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes

header-includes:
      - \usepackage{xr} \externaldocument[SI-]{SI_spatial_avg}
      - \input{head_maintext.sty}

tables: True
link-citations: True
urlcolor : blue
indent : True
csl: pnas.csl
bibliography: REF_CSS.bib
---

```{r setup_Paper, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
```

<!--Abstract-->
\noindent
\textbf{write abstract here.}
\linebreak

<!--keywords-->
\noindent
Synchrony | Tail-dependence | Serpentine grassland | Jasper Ridge
\linebreak

<!--# Introduction\label{Intro}-->
<!--Q1 : Estimating cross-species synchrony have any specific tail dependence, 
Q2: if yes, then what is the reason behind that?
Q3: Skewness change due to asymmetric taildep. as a new estimate of community variability?-->

<!--grassland and earlier work-->
\lettrine{U}{}nderstanding the mechanisms behind spatial and temporal 
dynamics has been a long-standing focus in community ecology [@gibson2009; @turner1989].
Interspecies interaction in a grassland 
community could be advantageous or
disadvantageous to either or both species involved, 
which led to mutualism or competition
as the case might be. These inter-species response were 
found to be affected by several biotic and abiotic factors 
including climate, fire, soil-nutrients, pollinators, micorrhizae,
grazing and other form of disturbances 
[@mcnaughton1983; @parton1995; @bardgett2008; @belsky1992; 
@collins1992; @olff1998; @fry2017; @potts2003; @hudewenz2012;
@canals2003; @hallett2014].
Excessive pressure of any of those 
disturbances could alter the competitive balance in the community
and could shift the species-dominance affecting overall species richness and 
ecosystem-function. Most of the prior studies used, in general, the correlation 
or regression based approach to understand species-variability with other
species within a community (or with environmental variables) to establish causal
relationship and thereby lacking important information embedded in their joint
distribution structure.

<!--Limitation of usual approach and introduction of copula?-->
Common correlation measures, i. e., Pearson's correlation used in earlier studies, is based 
on the notions of linearity, normality and mean which are not appropriate when dealing 
with extreme events or for non-normal distributions which occur commonly in ecological
datasets [our BIVAN paper]. Though rank-based non-parametric dependence measures, Spearman's 
$\rho$ and Kendall's $\tau$, do not assume linearity and normality, but they do not attribute 
sufficient weight to the extreme values and the distribution tail. Thus a better approach to model
multivariate dependence structure is necessary. Concept of a copula is a widely used 
statistical tool in other fields to model dependence between 
multiple variables [@joe2014_dependence; @nelsen2006_copula; 
@sklar1959_theorem; @burnham2003_modelselection]. A "copula" can separate the 
information contained in a joint distribution of bivariate random variables 
into two distinct parts: the information contained in the marginal 
distributions (which says nothing about the dependence between the 
variables), and the remaining information (which is solely
about the dependence between the variables). From a data perspective, we can construct 
a bivariate copula from a set of two variables
$\{ x_i, y_i\}, i = 1, 2, \ldots , n$ after some ranking procedure is applied so that 
their marginal distributions become uniform. If $\{ \bar{u_i}, \bar{v_i}\}$ are the
corresponding ranks of $\{ x_i, y_i\}$, then the "normalized rank" 
$\{ u_i, v_i\} = \{ \bar{u_i}, \bar{v_i}\} / (n+1)$
reveals the associated dependence in the bivariate copula when plotted in a unit square. 
Through suitable choice from a wide varity of families and parameters, one can control 
this dependence structure of a bivariate copula which shows over which parts of the 
distributions the variables are more strongly associated. This information
is implied in their degree of "tail-dependence" which can be viewed as stronger 
association between two random variables in their extreme values. If smaller values of 
two positively associated variables are more strongly correlated than are larger values, 
the variables are said to have stronger lower- or left-tail dependence than right- or 
upper-tail dependence (ADD SI FIG); and vice versa if higher values are more strongly
associated than smaller ones (ADD SI FIG). A copula might have symmetric 
(with zero tail-dependence in both tail) or asymmetric (finite tail-dependence in either
or both tail) feature depending on its family and parameters [@nelsen2006_copula].
Recently few studies have shown implication of copula in modelling spatial dependence of 
environental variables [@li2013; @serinaldi2008; 
@goswami2018; @she2018], but as of our knowledge,
it remains unexplored in studying interactive relationships between multiple
species and their responses with several biotic and/ or abiotoic factors among a community.
In this paper, we will preach for the potential of using "copula" approach to analyze
species' response within a community in understanding variation in species-richness
over a temporal scale. 

<!--why copula (and tail-dep.) is important in ecology, specially on grassland? : 
1) direction of competetion
2) leibig's law
3) global climate change : ongoing invasion and sp. extinction : pollinators?
-->
The relevance of copula with its unique tail-dependence feature is that it can be applied
to model synchronous (or compensatory) response between two or more ecological variables. 
Spatial synchrony [@liebhold2004; @walter2017; @sheppard2016] between 
metapopulations, is recently investigated using copula (BIVAN paper),
but there has been no report using copula approach in multi-species dynamics.
Here we will summarize how copula could be useful in ecology, specially in grassland community.
First, in case of compensatory dynamics between any two species over years, tail dependence
measurement using a bivariate copula not only shows species are negatively correlated but 
in addition could indicate which one is superior competitor between them. Secondly, earlier 
studies showed the importance of limiting resources (soil nutrients, N, P)
(Leibig's law) in the grassland productivity [@reynolds2003; @de2004; @john2007; @fay2015]. 
Therefore, a community showing overall
lower tail-dependence throughout years could suggest that the species-abundance on average
was limited as a particular resource became rare in that field and experimental addition of 
soil-fertilizer could improve their productivity (showing increase in upper-tail dependence)
after years. As multiple species in a grassland community could show limited growth
with respect to multiple nutrients and to different extent, therefore adding nutrient 
may increase the net primary productivity but could also reduce niche-dimensionality 
and increase compositional turnover [@harpole2016]. Using pairwise copula approach for
modelling a multivariate dependence structure offers more insight to understand
different species' response towards their drivers separately. Third, long-term and short-term
experiments established that change in global climate
(temperature, elevated CO$_{2}$ level, N-deposition) induced changes in species-composition 
in a grassland community 
[@cowles2016; @garcia2018; @harrison2015; @eisenhauer2013; @zavaleta2003]. 
As a result, some species could show high abundance with elevated
level of weather variables (upper-tail dependence) over years, but they also could suppress
and even lead to extinction of their weak competitors. Thus change in climate or 
resource has been proved to be a threat to the sustainability of native species 
in a grassland where invasive species replace them in their "favorable" scope on
space or time-scale [@white2001; @sm2001; @thomason2017].
This scenario could also evoke a potential threat for the pollinator 
community (e.g. butterfly) due to their habitat loss [@mclaughlin2002; @thomas2004].
Monitoring tail-dependence of species-above ground biomass (or percent cover) of 
a plant with respect to any of biotic and/ or abiotic factors could be beneficial
against this threat of biodiversity loss in a grassland eco-system. 

<!--community variability in view of skewness change?-->
Overall asymmetric tail-dependence (either lower or upper) bears a major implication
in determining community-variability and thus in turn stability of its compositional balance. 
A lower (or upper) tail dependence in community structure arises due to simultaneous rare
(or abundant) occurrence of the species and thus the "spikiness" of their individual 
time series become more or less synchronous. This synchronous response of multiple species
within a community generates a "skewed" distribution of total biomass over 
years. Skewness of a synchronous community could be significantly different from a community 
which either shows no interspecific tail-dependence or showing more compensatory dynamics 
so that their tail dependence cancel out. Significantly lower skewness (arising from lower 
tail dependence) has not only affected net primary productivity of a community 
but could also make some native species disappear from the system. 
Thus tracking skewness of a community over a period of years could serve
beneficial to better strategize conservation and management of native species. 
In terms of stability assessment, a community could be considered as less stable 
if the absolute value of skewness of total biomass time series is relatively high. <!-- am I right for this last sentence?-->

\begin{tcolorbox}[width=\columnwidth,colback={blue!20},arc=0mm,colframe=blue!50!black,coltext=blue!50!black]
Significance

Write significance here.
\end{tcolorbox}  

<!--Why choose jasper ridge and hays as study system?, what is our objective?-->
We consider two grassland ecosystem for our study. First one is Jasper Ridge Biological
Preserve (JRBP), a biodiversity hotspot in californian serpentine grassland. 
Long term observations in JRBP showed the serpentine grassland vary widely in its 
species composition both in temporal and spatial scale [@hobbs2007; @zhu2016]. 
Climate change and invasion by 
exotic species in JRBP had detrimental effect on native species shifting its dominance 
from one species to another
and thereby posing habitat loss of their
dependent species leading to extinction [@stromberg2007]. 
For example, bay-checkerspot butterfly 
\textit{Euphydryas editha bayensis} is now locally extinct at Jasper Ridge 
possibly due to declining abundance of its principal larval 
host plant \textit{Plantago erecta}
in competition with invasive species like \textit{Bromus hordeaceus}.
The second dataset we choose to invetigate is long-term mapping of 
individual herbaceous plants on permanent quadrats in a mixed grass 
priarie of Hays, Kansas. Hays grassland also played major role in investigating 
plants' demographic process, their co-existence and extreme climatic effect on annual productivity 
in the past studies [@fair1999demography; @adler2006climate; @lauenroth2008demography; @heisler2009contingent].
Both datasets have been served as experimental model system to 
study ecological hypothesis since long and proved to be effective
in decision making from conservation and management perspective. 

Our objective is to find the 
synchronous (for positively correlated variables)
or compensatory (for negatively correlated variables) dynamics in 
the serpentine grassland of JRBP and prairie of Hays in Kansas 
using "tail-dependence" feature of copula.
Implementing a model selection approach as well as nonparametric analysis, we 
try to find is there any asymmetric tail-dependence in those grassland community and
if there is, then what could we interpret about inter-species response from this
asymmetric tail-dependence? We also suggest an alternative measure of 
community variability (henceforth stability) in terms of quantifying skewness of 
total biomass distribution over years.

# Results\label{Results}

We present the results from the study carried out with the
spatio-temporal percent cover data (for 216 plots over 36 years, 1983-2018)
of several species found in the serpentine grassland
of Jasper Ridge Biological
Preserve (JRBP), located near Stanford
University in San Mateo County, California 
(for details see Appendix \ref{SI-Data}). We here analyzed
our findings from a model-selection based 
copula approach, but we also verified the overall finding using
alternative nonparametric analysis (Appendix \ref{SI-compare_methods}).

```{r read_res,echo=F}
RES_jrg_spaceavg<-readRDS("./Results/jrg_results/RES_jrg_spaceavg.RDS")
#res_skew<-readRDS("./Results/jrg_results/skewness_results/all_sp_39/skewness_jrg_spaceavg_spearman_common_sp_39.RDS")
```

<!--## Results from spatial averaging showing temporal dynamics \label{Results_spaceavg}-->
For a given year, we averaged percent cover data for each 
of 15 screened species over 216 plots. 
Tail-dependence analysis (see Appendix \ref{SI-Methods} for details)
was carried out with pairwise copula
made of those spatially averaged percent cover time-series.
The 19 $\times$ 19 matrix [15 plant species :
Table \ref{SI-tab_spinfo}
and 4 weather variables : total precipitation, average, 
maximum and minimum temperature 
for growing season (September-April)] 
in Fig. \ref{fig_LTmUT_spaceavg_jrg}
showed tail-dependence difference (lower minus upper) plotted 
for each pair-wise copula through 36 years (1983 - 2018). 
Spatially averaged data could only show a prominent
tail-dependence results in
our case when there is stronger correlation for some unusually 
good or bad years for both species' percent-cover considered.
We found considerable amount of interspecific competetion
and co-existence in the serpentine community. Here we will 
discuss a few exemplary cases. For instance,
when invasive species \textit{Bromus hordeaceus} were more abundant
throughout the plots over certain years,
native species \textit{Plantago Erecta} could not do well 
and became rare from the
system for those years, but when \textit{Bromus} was scarce in the
system \textit{Plantago} showed a varied abundance. 
This phenomenon indicated the causal relationship that
\textit{Bromus} appeared to be a stronger competitor
limiting \textit{Plantago}'s growth synchronously [represented in 
Fig. \ref{fig_LTmUT_spaceavg_jrg}
by the blue cell indicating 
stronger upper tail-dependence between yearly percent cover of
\textit{Bromus} (sp6 in row) and 
\textit{Plantago} (sp31 in column, plotted as negative
percent cover, so that the copula showed positive correlation,
see Fig. \ref{SI-fig_jrg_spaceavg_invasion} for details].
The interaction was conceptually same
as shown by the red cell with lower tail-dependence in 
Fig \ref{fig_LTmUT_spaceavg_jrg}(sp31 in row and
sp6 in column).
Similar finding was also observed earlier
by Hobbs et al.[@hobbs1991] where several years of above-average rainfall 
led to an abundance in \textit{Bromus hordeaceus} and shifted the dominance 
among native annuals from \textit{Plantago Erecta} to \textit{Lasthenia Californica}.

Strong co-existence with similar dynamics was found (both abundant simultaneosly 
in certain years) between two native annuals \textit{Hemizonia congesta}
and \textit{Lotus wrangelianus} (Fig. \ref{fig_LTmUT_spaceavg_jrg}, blue cell
with upper tail dependence between sp17, sp24) where the two species were 
abundant synchronously. When \textit{Hemizonia} and 
\textit{Lotus} were not dominantly prolific in the system, they did not 
show a strong correlation 
and another weak competitor \textit{Microseris douglasii} got a chance to flourish.
When those two species (\textit{Hemizonia} and \textit{Lotus}) were both 
abundant in their "good" years (specifically 3 successive years from 2016-2018,
(Fig. \ref{fig_LTmUT_spaceavg_jrg}), native annual \textit{Microseris}
(sp28) became uncommon and appeared as a weak competitor towards their joint
competitive pressure (sp17 and sp24 in rows with sp28 in column, for details
see Fig. \ref{SI-fig_jrg_spaceavg_competition}). If this trend continues, then it could be 
a potential threat for existence of native species \textit{Microseris douglasii}. 
This type of "diffuse competition" intrigued many experimental studies in the past 
to better understand plant competition and facilitation in serpentine grassland 
[@harrison2011] and in our study we showed tal-dependence could be a measure
of inter-specific competition. Another multispecies interaction was native annual 
\textit{Vulpia microstachys} (sp42) 
found to show stronger lower tail dependence (positive correlation with
simultaneous low-abundance) with other two species \textit{Layia Platyglossa} (sp19)
and \textit{Castilleja densiflora} (sp30). 

<!--Results including growing season ppt. effect on jrg spatial avg data-- 
Interestingly, we found
that annual and perennial showed different tolerance towards total precipitation 
for the growing season. For example, some extreme dry years [e.g. 2012 or 2014's 
drought [@griffin2014]]
with little precipitation made annual 
forb \textit{Castilleja densiflora} (sp30) uncommon whereas perennial forb 
\textit{Brodiaea spp.} (sp5)
became abundant at those years (see Fig. \ref{SI-fig_spaceavg_drought}). 
Earlier study [@pitt1978] showed years with "suboptimal" rainfall had relatively higher cover
of perennial forbs whereas years with "favorable" rainfall patterns promoted earlier
germination and better growth for annual grasses. Another perennial 
native bunchgrass \textit{Nassella pulchra} (sp36) became rare in the community with
low average temperature (considered as sp44)
during growing season for two successive years, showing lower tail-dependence.--> 

<!--Results from skewness as a measure of community variability-->
Major lower tail dependence than upper was observed in interspecific interactions 
as visually evident from Fig. \ref{fig_LTmUT_spaceavg_jrg}. We counted the number of positively correlated 
cells among $15 \times 15$ plant species interaction matrix and classified as $nL$
(showing more lower tail-dependence) and $nU$ (showing more upper tail dependence).
These numbers (or more specifically their associated asymmetric tail-dependence
values) play a significant contributors to the skewness of total community biomass
distribution over years. For each of the 36 years (1983-2018), we computed total percent cover
(averaged throughout the plots) of all species 
(except LASP : 2014-2018 data reported only, need clarification from Lauren, Bare Ground).
We measured skewness 
of this aggregated percent cover time-series and compared against skewnesses of $10000$
spearman-correlation preserving "Normal" surrogate time-series generated by a resampling approach
(for details see Appendix \ref{SI-skew_ncsurrog}). We found remarkable change in 
skewness (Fig. \ref{fig_skewness}) and that was marginally significant as compared 
to their "Normal" analog. As skewness is related to the "spikiness" of the time-series
concerned, lower skewness means the community could have a potential 
threat for some of its native species which could crash simultaneously below a threshold.

\begin{figure*}[!h]
\begin{center}
\includegraphics[width=12cm]{./Results/jrg_results/LTmUT_jrg_spaceavg.pdf} 
\caption{Model averaged lower tail - upper tail dependence matrix for 15 most common plant species, 
details were discussed in \nameref{Results}. 
nL and nU are the count of cells (considered for positive correlation only)
with positive and negative tail-dependence difference 
values for 15$\times$15 matrix; 
nL$>$nU (majority of red cells) indicates major lower tail dependence
in sp-sp interaction matrix.
Yellow cells indicate independent relationship. \label{fig_LTmUT_spaceavg_jrg}}
\end{center}
\end{figure*}

\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15cm]{./Results/hays_results/LTmUT_hays_spaceavg.pdf} 
\caption{Model averaged lower tail - upper tail dependence matrix for 20 most common plant species, 
details were discussed in \nameref{Results}. 
nL and nU are the count of cells (considered for positive correlation only)
with positive and negative tail-dependence difference 
values for 15$\times$15 matrix; 
nL$>$nU (majority of red cells) indicates major lower tail dependence
in sp-sp interaction matrix.
Yellow cells indicate independent relationship. \label{fig_LTmUT_spaceavg_hays}}
\end{center}
\end{figure*}




# Discussion



### \textbf{For Data and Methods, see Appendix \ref{SI-Data} and \ref{SI-Methods} in Supporting material, respectively.}





\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{1pt}
\noindent









