---
title: "Species relationships in the extremes and their specific influence on community stability"
author:
 - Shyamolina Ghosh\textsuperscript{1}, Daniel C. Reuman\textsuperscript{1,2,\dag}
 - \textsuperscript{1}Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA; \textsuperscript{2}Laboratory of Populations, Rockefeller University, 1230 York Ave., New York, NY 10065, USA
date: '\textsuperscript{\dag}Correspondence: Daniel C. Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084'
fontsize: 11pt
geometry: "left=1 in,right=1 in,top=1 in,bottom=1 in"

output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes

header-includes:
      - \usepackage{xr} \externaldocument[SI-]{SI_spatial_avg}
      - \input{head_maintext.sty}

mainfont: Times New Roman 
tables: True
link-citations: True
urlcolor : blue
indent : True
csl: ecology-letters.csl
bibliography: REF_CSS.bib
---

<!--***DAN: store alternative titles here, decide on one closer to the end
1) Skewness ratio: a complimentary measure for variance ratio and community stability
2) A complete description of synchrony and compensatory dynamics yields a better understanding of community stability
3) Asymmetric tail association, skewness and community stability
4) Asymmetric tail association, skewness and community variability 
5) Asymmetric tail association affects community stability
6) Species relationships influence community stability in more ways than classical approaches reveal. This is OK, but
may make people think of it as a methods paper?
7) Through thick and thin: species relationships in extreme times and their specific influence on community stability 
8) Species relationships in extreme times and their specific influence on community stability
-->

<!--***DAN: Shyamolina, please go ahead and format the title page appropriately, making the proper affiliations
appear and so on
***Shya: affiliation added, I remember Prof. Foster and Prof. Bever listened to your talk and gave input, should not I include them as co-authors?
***DAN: Good question. I don't think their contribution rises to the level of authorship yet, but I've added them to the 
acknowledgments, and I think we should send them a draft at some point and see if they want to contribute more. How does that 
sound to you?
***Shya: Fine for me.-->

```{r setup_Paper, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
```

<!--Abstract-->
\newpage

\noindent \textbf{ABSTRACT.} 

<!--keywords-->
\noindent \textbf{\textit{Keywords:}} variance ratio, community-stability, synchrony, tail-association, grassland, skewness


\newpage

# Introduction\label{Introduction}

Understanding the extent to which species interactions in an ecological 
community help stabilize or destabilize the ecosystem as a whole is a topic that has fascinated ecologists 
for decades [@macarthur1955fluctuations; @paine1969note; @grimm1992application; @pimm1984complexity; @gonzalez2009causes]. 
An important early insight was that an aggregate 
community property such as the total biomass of all species can be relatively stable through time, even 
while the dynamics of individual species are highly variable, if different species exhibit offsetting fluctuations.
Such fluctuations are often referred to as *compensatory dynamics*, because decreases in species abundances are compensated
for by simultaneous increases in other species [@frost1995species;@bai2004ecosystem;@hallett2014]. When species instead exhibit 
positively correlated fluctuations through time, it tends to cause increased variability through time in the community 
aggregate property [@keitt2008coherent;@loreau2008species;@ma2017climate]. This is called *synchrony* because 
species dynamics are partly synchronized in their fluctuations. Thus an important part of what determines 
the stability/variability of an aggregate property such as total biomass is the nature
of relationships between the dynamics of different species. 

Historically, how species relationships influence community stability, including questions of compensatory 
and synchronous dynamics, have been investigated principally by examining the coefficient of variation through time
of the aggregate community property, together with covariances between species population abundance time series.
For instance, if $x_i(t)$ are data representing the population or biomass abundances of species $i=1,\ldots,N$ at the sampling times
$t=1,\ldots,T$, and if $\mu_i=\mean(x_i)$, $v_{ii}=\var(x_i)$, and $v_{ij}=\cov(x_i,x_j)$, community variability
has commonly been quantified using the squared coefficient of variation of $\xtot(t)=\sum_i x_i(t)$, i.e., 
$\CVcomsq=\var(\xtot)/\left(\mean(\xtot)\right)^2=\left(\sum_{i,j} v_{ij}\right)/\left(\sum_i \mu_i\right)^2$.
And this formula obviously connects to species relationships as measured via 
the covariances, $v_{ij}$, which appear directly in the formula. Denoting by 
$\CVindsq=\left(\sum_i v_{ii}\right)/\left(\sum_i \mu_i\right)^2$ the value that $\CVcomsq$ would take if 
the dynamics of each species were independent of the dynamics of all other species (so $v_{ij}=0$ for all $i \neq j$), the 
classic *variance ratio* was defined [@schluter1984variance;@peterson1975] as 
$\phi_{\text{cv}}=\CVcomsq/\CVindsq=\var(\xtot)/\left(\sum_i v_{ii}\right)=\left(\sum_{i,j} v_{ij}\right)/\left(\sum_i v_{ii}\right)$. 
Thus $\CVcomsq=\phi_{\text{cv}} \CVindsq$. Because $\CVindsq$ is interpreted as what community variability would be, without species interactions,
$\phi_{\text{cv}}$ is the factor by which species interactions inflate or decrease community variability.
The variance ratio has therefore commonly been used [@houlahan2007;@winfree2009;@hallett2014] as an index of whether dynamics 
are synchronous or compensatory. If $\phi_{\text{cv}}>1$, the interpretation has been that dynamics are, on balance,
synchronous because then $\CVcomsq>\CVindsq$, i.e., the total community abundance $\xtot$ is more variable than
it would be if species dynamics were independent. Conversely, if $\phi_{\text{cv}}<1$, the interpretation has been
that dynamics are, on balance, compensatory. 

Timescale-specific extensions of the variance ratio [@vasseur2007;
@brown2016;@zhao2020] have revealed that it is possible for compensatory 
dynamics to occur on some timescales while 
synchronous dynamics simultaneously occur on other timescales, and other improvements and alternatives to the variance ratio 
have also been proposed [@loreau2008species]. However, $\phi_{\text{cv}}$ is still commonly used; 
furthermore, as we will show, how species relationships influence community stability is likely to be 
more nuanced than is revealed by either the classic variance ratio approach or
by any existing extensions of which we are aware. To 
begin to illustrate the concepts we will explore in this 
study, and to understand what has been omitted from earlier approaches, we consider two artificially
generated community dynamics datasets (Fig. \ref{fig_pedag_cv2_skw}A, B; data generated as in SI section \ref{SI-PedagogFigMethods}).
Both model communities consist of $N=20$ species, and data were generated so that species marginal distributions 
(and therefore the means, $\mu_i$, and variances, $v_{ii}$, 
through time) were the same, up to sampling variance, in both scenarios. Likewise, data were 
generated so that pairwise species covariances, $v_{ij}$ for $i \neq j$,
were essentially the same for Fig. \ref{fig_pedag_cv2_skw}A and B. Thus both $\CVcomsq$ and $\CVindsq$, which 
depend only on the $\mu_i$ and $v_{ij}$,
were the same in both scenarios, and so $\phi_{\text{cv}}$ was also the same. The Loreau-de Mazancourt variance ratio
[@loreau2008species], which itself depends only on the $v_{ij}$, was also the same, up to sampling variance, in the two scenarios 
(Fig. \ref{fig_pedag_cv2_skw}A, B). Classical approaches such as these do not distinguish between our two scenarios, neither in the nature
of species interrelationships nor in the effects of interrelationships on community aggregate dynamics $\xtot(t)$.

Nevertheless, species relationships and resulting aggregate dynamics differed in important and related ways in our two scenarios. In both scenarios,
the twenty species could be separated into two groups of ten species (red and black lines on Fig. \ref{fig_pedag_cv2_skw}A, B), 
and the dynamics of species in different groups were strongly compensatory. However, in scenario 1 (Fig. \ref{fig_pedag_cv2_skw}A), 
species in the same group exhibited strongly synchronous dynamics when rare and much less synchronous dynamics when 
common; whereas in scenario 2 (Fig. \ref{fig_pedag_cv2_skw}B) species in the same group exhibited strongly synchronous dynamics 
when common and much less synchronous dynamics when rare. Dynamics of this kind could occur ecologically, for instance, in scenario 1, 
if red species in Fig. \ref{fig_pedag_cv2_skw}A were sensitive to an environmental factor such as precipitation, but only when it 
takes values below a threshold. So all the red species are controlled by the same factor, precipitation, when it is below the threshold,
and hence are synchronous when rare. 
When precipitation is above the threshold, the red species are each sensitive, instead, to other, distinct factors, rendering them asynchronous
when common. Black species 
of Fig. \ref{fig_pedag_cv2_skw}A may also be sensitive to precipitation, but in a reverse manner, being all simultaneously 
harmed by precipitation when it is
above the threshold, and unaffected by precipitation below the threshold. Resulting community dynamics 
(Fig. \ref{fig_pedag_cv2_skw}C-F) were quite different in our two scenarios because of the distinct species interrelationships.
Total community abundance was left-skewed, frequently crashing to low values, for scenario 1; but was right skewed, frequently
exploding to high values, in scenario 2. It is not difficult to imagine cascading ecological ramifications or 
potential applied significance of the distinction our example illustrates. For instance, grazers subsisting on a grass mixture that 
exhibits occasional crashes may experience different growth prospects than grazers consuming a grass mixture that
instead shows occasional explosions of abundance. 
Thus species interrelationships and their effects on community aggregate dynamics may differ in potentially ecologically
meaningful ways which are nevertheless not detected by classical approaches. 
It is the reliance of earlier approaches on variances, covariances and related linear tools
that causes those approaches to be unable to register the differences
we have illustrated here.

This study is centered around three suggestions about how research on these topics should develop in the future;
and we take some initial steps toward implementing these suggestions. First, descriptions of species 
relationships that are more nuanced than the standard 
covariance/correlation approaches ought to be developed and applied.
Fig. \ref{fig_pedag_cv2_skw} suggests that measures might be
useful that isolate the extent to which species abundances are more related to each other when the species
are common, or rare.
Second, statistical descriptions of community 
stability/variability that are more nuanced than the coefficient of variation ought to be developed 
and used. The results of Fig. \ref{fig_pedag_cv2_skw} suggest that measures related to skewness may
be useful. Third, we ought to then study how species relationships
influence community stability/variability using the nuanced descriptors. 

The core ideas of this study relate to rare and extreme events, and specifically to how the relationships between the constituents of 
an aggregate quantity change during such events. There are strong conceptual and methodological connections to
risk pricing in finance. A portfolio of investments is supposed to reduce risk if the constituent investments 
are negatively correlated or uncorrelated. But if these constituents become positively correlated, it can instead accentuate risk. 
Some observers have blamed the 2008 financial crisis on a trend during which many analysts estimated correlations between housing default
risks based on data from normal times, not taking into account that in a crisis, default risks become much more correlated.
In such times, financial products based on aggregating mortgages become riskier due to the increased correlation. 
A single modelling approach [@li2000] was apparently widely used [@salmon2009]. The relationship between two 
real quantities, be they investment risks or species populations, can probably never be completely captured by a single number
such as a covariance [@cheung2013], nor considered to be constant through time. Just as mortgage defaults which were typically uncorrelated 
became correlated when the macroeconomic situation worsened, it seems likely that abundances of two species in a community may
be more or less correlated with each other under different circumstances. Therefore classical approaches that characterize relationships
between species using covariances may be neglecting something important. In particular, correlations between species 
may change as global climate change progresses, or when extreme climatic events impact an 
ecosystem [@Ghosh_extinction]. Some financial analysts 
have learned these lessons and have begun developing the necessary mathematics for more appropriate models [@cheung2013]. Ecologists 
can benefit from application of the same methods, and we hope this study helps begin that process.

The goals of this study are as follows, and parallel the research suggestions listed above. (G1) We will develop 
and apply an approach to studying the relationships
between species that complements classic approaches relying on covariances. (G2) We will also develop and apply an
approach to studying aggregate community stability/variability that complements the traditional use of the 
coefficient of variation. Finally, (G3) we will study how species interrelationships influence community stability/variability
using the new approaches. Our results provide a new take on classic questions about the influence of synchronous or 
compensatory dynamics on community stability. We will use theory, as well as data from two iconic, long-term grassland studies:
plant percent cover data from long-term quadrats in mixed-grass prairie located in Hays, Kansas; and annual 
species composition data from the Konza Prairie Long-Term Ecological Research (LTER) site, also in Kansas.
Synchronous versus compensatory dynamics in communities, and their influence on community stability, has most commonly
been studied in plankton systems [@downing2008;@keitt2008coherent;@brown2016] and in grasslands 
[@bai2004ecosystem;@houlahan2007;@hallett2014]. Evidence so far suggests that synchronous dynamics may be dominant
and compensatory dynamics rare in plankton systems [@Vasseur2014]. On the other hand, previous work suggests that
compensatory dynamic may be at least somewhat more common
in grasslands [@hallett2014;@zhao2020]. Though some studies have also found compensatory dynamics to be rare in
grasslands [@houlahan2007;@valone2008], others have found general evidence for compensatory dynamics [@bai2004ecosystem;@hector2010] 
or have found compensatory dynamics selectively, in some contexts [@grman2010;@hallett2014]. 
One outcome of our empirical work is a suggestion that compensatory dynamics in grasslands may be less 
common than was previously appreciated, when one adopts a broader conceptual view of what constitutes synchronous versus 
compensatory dynamics. Overall, our work seeks to establish the beginnings of a new and more flexible paradigm for
conceptualizing synchronous and compensatory dynamics and their influence on community stability.

<!--***DAN: Not sure how relevant this stuff is, but keeping it around for now, partly for the references.
There are reasons to focus on skewness to better understand the community-stability 
in this rapidly changing world. Earlier studies [@smith2011ecological; @hoover2014resistance; 
@felton2017integrating] showed climatic events that are extreme and 
sometimes spatially well-spread impact the net productivity of community [@ciais2005europe],
community assembly [@thibault2008impact] and even induce a shift in 
ecotone boundaries [@allen1998drought]. Not only the mean and variance, the symmetry of 
temporal distribution for such climatic signals is also changing [@ummenhofer2017extreme]. This 
change in symmetry results into change in the skewness that can act on the community as well 
as on ecosystem level. Another reason is like variance changing skewness had been 
used earlier as an indicator for changing ecosystem (i.e. for an early 
warning signal, regime shift, etc.) [@guttal2008changing; @guttal2009spatial; 
@dakos2010spatial; @dakos2012methods]. But to our knowledge, there is a gap to incorporate 
this skewness into direct community-stability metric.-->

<!--***DAN: Keeping the following text, may end up included in the Discussion.
A recent study [@Ghosh_copula]
shows tail associations between local quantities affects the 
skewness of spatially averaged
productivity. Likewise, we anticipate the tail-association between 
individual time-series could be responsible for changing skewness on 
an aggregate level. Tail association means a stronger correlation between two 
variables at their extremes than moderate values. If two random variables (say, 
time series for two individual species from a community) are more 
associated in their smaller values than higher then they are having 
stronger 'lower-tail' association (Fig. \ref{fig_pedag_taildep}C) indicating 
both species are going synchronously rare than abundant 
(as shown in Fig. \ref{fig_pedag_cv2_skw}A). Similarly, 
if variables are more associated in their higher values than smaller
then they are having stronger 'upper-tail' association (Fig. \ref{fig_pedag_taildep}D)
indicating both species are going synchronously abundant than rare 
(as shown in Fig. \ref{fig_pedag_cv2_skw}B). A finite 
difference between lower- and upper-tail association indicates 'tail-asymmetry' 
between interspecific response over the years. Tail-asymmetry is known to
have significant effect on population-extinction risk [ERC_ecosphere]
and hence could be related to stability on the population, community and even ecosystem level.
For example, the lower tail association could cause the community 
likely to crash at once - making the community less stable.-->

<!--p6: Our objectives:
(Q1) Could we find an overall tail association (L/U) in a community which arises because of  
synchronous/ compensatory species interaction? [answer by pscor]
(Q2) If we do, then how does tail-association pattern explain the underlying mechanism 
affecting the overall skewness of the community? More specifically, could this
assymetric sp. interaction explain the community stability which we might miss 
with the classic variance ratio method? [answer: Fig 1b hidden box revealed]
We will answer these Qs with two grassland community datasets: Hays & Konza.-->

<!--***DAN: These framings of the objectives are too technical for the reader to understand right now,
keeping for possible use at end of Theory section.
In this paper, we will explore the potential for asymmetric tail associations 
to assess nonlinear interactions among species in a community and to 
unravel any hidden mechanism - not detected by the usual variance ratio approach -
for stabilizing (or destabilizing) a community. Specifically, we will address 
the following sets of questions: (Q1) In a community, could we find 
asymmetric tail associations (lower and/ or upper) that arise because of 
synchronous/ compensatory interaction between species? Do those interactions 
make the community overall lower-tail or overall upper-tail asymmetric? 
(Q2) If we do, then how does the tail association pattern 
explain the underlying mechanism affecting the overall 
skewness of the community? To be specific, we seek for 
a route map to predict stability for an overall tail-asymmetric 
community, based on a 'skewness ratio' approach. For (Q1), we use partial 
Spearman correlation measures [@Ghosh_copula; ERC_ecosphere] focusing 
on the extremes of pairwise species interaction. For (Q2) we disentangle
the overall contribution in variance-/ skewness-ratio 
into two distinct parts: one from the usual correlation and the 
other from the tail association. We tested our theoretical 
framework for assessing stability measure and detecting tail-asymmetry for 
a community with two long-term grassland datasets: basal cover data of vascular 
plant species from Hays and plant species composition data from Konza prairie. 
To our knowledge, though the potential of tail association in population 
ecology [@Ghosh_copula; ERC_ecosphere] has been recently studied, using it 
on the community level to assessing variability-stability relationship remains 
unexplored. Our essential goal is to advance the idea that tail
associations may also be important in this context. Our questions are 
also an important step for predicting net productivity, 
community reorganization, etc. through monitoring skewness over years.-->

<!--idea: 
- can we make specific outcome oriented 
community by choosing species or a cluster of sp.?
- could change in skewness/ phi_s indicate species reordering 
similar to that of early warning signal 
used for regime shift?-->

<!--Tables
S1: splist for hays
S2: splist for knz
-->

# Theory \label{Theory}

\noindent In this section we illustrate theoretically how relationships between
constituent species of a community, beyond just pairwise covariances/correlations of the species
time series, can influence the variability of the 
total community abundance. The setup for the theory is as follows. As in the Introduction, let $x_i(t)$ be an abundance 
measure for species $i=1,\ldots,N$ at time $t=1,\ldots,T$. We use $N=20$ and $T=100000$ in the examples here. 
We assume the multivariate random variables $x(t)=(x_1(t),\ldots,x_N(t))$ are independent
and identically distributed for distinct times, $t$. It should be straightforward to replace
this assumption with an appropriate assumption of stationarity and ergodicity, with no real change
to the theoretical development, but we adopt the stronger assumption for simplicity. 
To describe relationships between species, 
classic approaches use the covariances
$v_{ij}=\cov(x_i,x_j)$, and related quantities. We will instead consider how 
$x_i$ and $x_j$ are related to each other in their tails.
To describe community variability, classic
approaches consider the variance of 
$\xtot=\sum_i x_i$, and related quantities. But this is only a summary of the 
distribution $\xtot$. We will instead consider the 
whole distribution of $\xtot$, and its skewness. 

We present four scenarios (1-4 below). For all scenarios, the marginal distributions $x_i$ and
the covariances $v_{ij}$ are the same, as are the quantities $\var(\xtot)$, $\CVcomsq$, 
and $\phi_{\text{cv}}$. Thus 
our scenarios are indistinguishable from the standpoint of classic approaches. But our scenarios are
nevertheless ecologically distinct. As we will see, the total abundance $\xtot$ achieves 
more extreme values under some scenarios than others. Hence our scenarios represent communities showing 
different characteristics of stability/variability. 
Scenarios are intentionally as simple as
possible, with complexities of population dynamics such as autocorrelation not included, so that 
new ideas can be clearly described. But the ideas here apply to real communities, as will be demonstrated. 
We also compare our four scenario to a reference scenario, called the *comonotonic* scenario, for which 
the $x_i$ are deterministically related via perfect positive relationships. 
This is a scenario of maximal covariance between species
and a maximal value of $\var(\xtot)$ [@cheung2013]; it is a scenario of maximal community instability, given
fixed species marginal distributions, $x_i$.

In scenario 1, which is a baseline for comparison, 
$(x_1,\ldots,x_N)$ is a multivariate normal distribution with mean $(0,\ldots,0)$ and covariance matrix
having $1$s along the diagonal and off-diagonal entries $0.6$. A sample from $(x_1,x_2)$ (Fig. \ref{fig_theory}A)
and marginal histograms (side panels of Fig. \ref{fig_theory}A) help illustrate the distribution. 
The standard deviation of $\xtot$ was about $`r round(sqrt(vartotX1),1)`$, and the skewness of $\xtot$
was close to $0$ (Fig. \ref{fig_theory}B). The probability of $\xtot$ exceeding a high threshold is given 
in Fig. \ref{fig_theory}C for a range of high thresholds. Likewise the probability of $\xtot$ falling below a
low threshold is in Fig. \ref{fig_theory}D for a range of low thresholds. These probabilities represent the 
possibility that $\xtot$ will surpass or fall below a threshold considered disasterous or unacceptable from an
applied or ecological viewpoint, and we refer to these as *disaster thresholds*.

For our other baseline scenario, the comonotonic scenario (scenario C), we assume $x_i=x_1$ for all $i$. 
This is a specific form of comonotonicity [@cheung2013]. The relationship between $x_1$ and $x_2$ is 
pictured in Fig.  \ref{fig_theory}E.
Because synchrony between the dynamics of individual species is perfect, $\sd(\xtot)$ is much larger
than for scenario 1 (Fig.  \ref{fig_theory}F). Probabilities of $\xtot$ exceeding a high disaster threshold or falling
below a low one are likewise larger (Fig.  \ref{fig_theory}G-H).

Scenarios 2-4 illustrate the ideas, previously little recognized in ecological research, 
of tail comonotonicity [@cheung2013] and tail associations [@Ghosh_extinction;@Ghosh_copula], and the consequences of tail
associations for $\xtot$ and the extreme values of $\xtot$. 
In scenario 2, as in all our scenarios, each $x_i$ is standard-normally distributed (Fig. \ref{fig_theory}I side panels),
but $(x_1,\ldots,x_N)$ is not a multivariate normal distribution. Instead, $(x_1,\ldots,x_N)$ was engineered so
that $x_i$ and $x_j$ are perfectly related when they exceed some threshold, but imperfectly related 
below the threshold (Fig. \ref{fig_theory}I; see SI section \ref{SI-TheoryMethods} for 
details of this and the other scenarios). The strength of association below the threshold was chosen 
to make $\cov(x_i,x_j)$ equal to the same value as in scenario 1, so $\sd(\xtot)$ was also
the same, up to sampling variation, as scenario 1 (Fig. \ref{fig_theory}J). (Recall that 
$\sd(\xtot)=\sqrt{\var(\xtot)}$ and $\var(\xtot)=\sum_{i,j} \cov(x_i,x_j)$.)
Comonotonicity in the upper tails produces right/positive skewness in $\xtot$ (Fig. \ref{fig_theory}J).
Skewness translates to elevated probabilities of $\xtot$ exceeding high disaster thresholds (Fig. \ref{fig_theory}K),
compared to scenario 1. The probability of exceeding sufficiently high disaster thresholds is, in fact, the
same as scenario C. So although scenarios 1 and 2 are indistinguishable to classic approaches, the probability 
of disaster in scenario 2 is much elevated, and is as high as it can be given the species marginal
distributions used. Although it does not appear unstable to classic approaches, the scenario 2 community
is maximally unstable with respect to high disaster thresholds.

Scenario 3 is symmetrically like scenario 2, with comonotonicity in the left tails instead of the right (Fig.  \ref{fig_theory}M-P).
Comonotonicity in the lower tails produces left/negative skewness in $\xtot$, 
which in turn translates to elevated probabilities that $\xtot$ will fall below low disaster thresholds (Fig. \ref{fig_theory}P).
For sufficiently low thresholds, the probability is the same as for scenario C, and is therefore maximal for the 
species marginal distributions used. 

Scenario 4 shows comonotonicity in both tails, and therefore probabilities of $\xtot$ exceeding high disaster thresholds or falling below
low disaster thresholds are both as high as for scenario C, and are maximal. Although 
scenarios 1-4 are indistinguishable to classic approaches, 
the probabilities of disaster are elevated, and are maximal given species marginal distributions. 
The communities in scenarios 2-4 are each maximally unstable in a different respect, although they are indistinguishable from
each other, and from the relatively stable scenario 1, to classic approaches. 

Our scanarios were constructed making use of mathematical results of @cheung2008 and @cheung2013. We refer the reader
to those papers for mathematical specifics, while summarizing some of their importance for ecological
applications here. Given one-dimensional cumulative distribution functions (CDFs) $F_1,\ldots,F_N$, the 
\emph{Fr\'{e}chet space} $\mathcal{R}(F_1,\ldots,F_N)$ is the set of all $N$-dimensional random vectors 
$(x_1,\ldots,x_N)$ with the $F_1,\ldots,F_N$ as their marginal distribution functions. Thus the $\text{Fr{\'e}chet}$ space
represents, in our modelling context, all possible interspecific relationships (different degrees and kinds of 
synchrony and compensatory dynamics between species) that can pertain, given fixed
species marginal distributions. So the question of what forms the distribution of $\xtot$ can take across the $\text{Fr\'{e}chet}$ space
is a precisely formulated version of the classic question of how species interrelationships 
such as synchrony and compensatory dynamics can influence community stability/variability.
It is well known that the sum
$\xtot$ exhibits maximal variance when the $x_i$ are comonotonic [@cheung2013]; this is the maximal-synchrony case.
@cheung2013 showed the converse, i.e., if $\var(\xtot)$ is maximal in the $\text{Fr\'{e}chet space}$, then 
$(x_1,\ldots,x_N)$ must be comonotonic (subject to some mild regularity assumptions about the $F_i$). 
The classical, variance ratio approach essentially makes sole use of
$\var(\xtot)$ to quantify community variance ($\CVcomsq$ is used, actually, but this is equivalent, since species 
marginals are fixed within the $\text{Fr\'{e}chet space}$). But @cheung2013 showed that, 
even if we consider only the sub-$\text{Fr\'{e}chet}$ space
consisting of random vectors $(x_1,\ldots,x_N)$ in $\mathcal{R}(F_1,\ldots,F_N)$ that additionally have 
$\var(\xtot)=S$ for some fixed $S$, there are random vectors that achieve extreme values just as extreme as the
comonotonic case (e.g., scenarios 2-3 in Fig. \ref{fig_theory}). The classic approach tells us nothing
about true community stability/variability, in the sense that if species relationships are tail associated,
the community can exhibit the same extreme behavior as the comonotonic case, while the classically important
quantity $\var(\xtot)$ gives the impression of a much stabler community.

# Methods\label{Methods} 

## Data

<!--Data prep, and common/intermediate/rare species-->
\noindent We used data from two grassland sites that have been monitored for decades. 
After initial processing, data from the Hays, Kansas site consisted of annual estimates of basal cover, 
averaged over 36 quadrats from which livestock were excluded, for each species present, for the years 1932-1972
[@adler2007long]. 
Data from Konza Prairie, after initial processing, consisted of annual canopy percent cover data by species for the years
1983-2018, averaged over 20 annually burned plots on tully soils which were not grazed by livestock [@Hartnett2019]. 
Additional details of the sites and initial data processing are in SI section \ref{SI-Data}.

Community data often include numerous rare species, including species which were only detected a few times over the 
duration of a study, and our data are no exception. Rare species complicate analyses of interspecific relationships 
and their effect on community variability because the relationship of a species with a rare species is difficult to accurately
assess. We therefore categorized species as "common" (present in one or more of the quadrats considered for at last 35 years),
"rare" (present for at most two years), or "intermediate" (other species). Rare and intermediate species were combined
into a single pseudo-species for analyses. Common species made up the overwhelming majority of both the Hays and Konza 
communities that we studied (Fig. \ref{SI-fig_spaceavg_timeseries}), so it is the variability of these species, and their
interspecific relationships, that mainly determine community stability. Common species are listed in Tables
\ref{SI-tab_spinfo_hays} and \ref{SI-tab_spinfo_knz}. Henceforth the terminology "species" signifies
both true species the psudo-species formed by combining rare and intermediate species.

## Statistical methods

<!--Normalized rank plots and partial spearman-->
\noindent Fig. \ref{fig_pedag_cv2_skw} and the Theory section
suggest that to accomplish goal G1 of the Introduction, we could apply measures of tail association between species, 
and asymmetries of tail association. So we here define such measures, based on the *partial Spearman correlation*
of @Ghosh_copula. Given $x_i(t)$ for $t=1,\ldots,T$, we begin by defining the
*normalized rank* $u_i(t)$, equal to the rank of $x_i(t)$ in the set $\{x_i(1),\ldots,x_i(T)\}$, 
divided by $T+1$. The smallest element of this set is here considered to have rank 1, and the largest element to 
have rank $T$. Plotting the normalized ranks $u_i(t)$ and $u_j(t)$ against each other for two positively associated species 
$x_i$ and $x_j$ provides a visual
indication of tail association and asymmetries in tail association (Fig. \ref{fig_pedag_taildep}), 
as described by @Ghosh_copula.
Given two bounds $0\leq b_l < b_u \leq 1$ ($l$ stands for "lower" and $u$ for "upper"), we define the lines $u_i + u_j=2b_l$ and $u_i+u_j=2b_u$,
which intersect the unit square $[0,1]\times[0,1]$ of the normalized rank plot. The partial Spearman
correlation associated with the band circumscribed by these bounds is
\begin{equation}
\cor_{b_l,b_u}(x_i,x_j) = \frac{\sum (u_i(t)-\mean(u_i))(u_j(t)-\mean(u_j))}{(T-1)\sqrt{\var(u_i) \var(u_j)}},\label{PartialSpearmanEq}
\end{equation}
\noindent where sample means and variances are computed over $t=1,\ldots,T$ but the summation in the 
above formula is only computed over $t$ such that $u_i(t) + u_j(t)>2b_l$ and $u_i(t) + u_j(t)<2b_u$.
The partial Spearman correlation is the component of the Spearman correlation which can be attributed to
the points in the band. For two positively associated variables, $\cor_{0,b_u}$ for $b_u \leq 0.5$ measures 
lower-tail association, and $\cor_{b_l,1}$ for $b_l \geq 0.5$ measures upper-tail association. 
The difference $\cor_{0,0.5}-\cor_{0.5,1}$ is then a way to measure asymmetry of tail
associations. Positive values of this statistic indicate stronger lower-tail association, and negative values 
indicate stronger upper-tail association. We henceforth use the shorthand $\cor_l$ for $\cor_{0,0.5}$ and 
$\cor_u$ for $\cor_{0.5,1}$. 
@Genest2007 recommend making inferences about dependence structures such as tail associations using 
rank-based approaches such as the tools we introduced here. 
@Ghosh_copula, @Ghosh_extinction and @Ghosh_aphidplankton provide additional information on
why rank-based approaches and the partial Spearman correlation are an appropriate statistical choice 
for purposes such as ours. 

<!--***DAN: I think we should handle the negatively associated species problem in another way, below, but 
keeping this text for now in case we need it later.
upper tail association (Fig. \ref{fig_pedag_taildep}D). Though the aforementioned notion of 
tail-asymmetry measure is constructed for positively correlated species pair keeping in mind, it 
is possible to extend the idea for negatively correlated species. When the 
species are competing (i.e. overall correlation is negative), we can take any one species time 
series as negative (so that now species are positively correlated) and 
make a similar plot as in Fig. \ref{fig_pedag_taildep}C or D (depending on which 
species we are reversing). Species for reversal can be chosen arbitrarily, 
hence only the magnitude, not the sign of tail-asymmetry 
$\cor_{l}-\cor_{u}$ is considered for competing species pair. Therefore, a finite 
value of $|\cor_{l}-\cor_{u}|$ would indicate asymmetric tail association 
between two corresponding species time series $\{ u_t, v_t\}$. In particular, 
the magnitude for compensatory and magnitude as well as the sign for synchronous 
(positive for lower tail and negative for upper tail) dynamics can answer the 
first part of (Q1). 
-->

<!--Aggregating to the community-->
We computed $\cor_l-\cor_u$ for all positively associated species pairs (those with positive Spearman correlation); negatively associated 
pairs were ignored. Fig. \ref{fig_pedag_cv2_skw} and the Theory section
suggest that the aggregate nature of tail associations of positively associated species pairs may influence
the distribution of $\xtot$, i.e., communities with a preponderance of upper-tail association between positively
associated species will have $\xtot$ with distinct distributional properties from communities
with a preponderance of lower-tail association. Community-aggregated tail association between positively 
associated species was measured by counting the number, $n_L$, of values
$\cor_l(x_i,x_j)-\cor_u(x_i,x_j)$, for $i \neq j$, that were positive, representing stronger lower- than
upper-tail association between positively associated species; as well as the  number, $n_U$, of values 
that were negative, representing stronger upper- than lower-tail association.
We also computed the sum, $\Atot$, of $\cor_l(x_i,x_j)-\cor_u(x_i,x_j)$
across all positively associated species pairs, $i,j$. This was positive if lower-tail 
association was, in aggregate across the community, stronger than upper-tail association, and
negative if the reverse. We refer to $\Atot$ as the *total community tail association*
of positively associated species pairs. 
Note that the word "positive" applies in three senses, here, all with distinct meanings: positive association 
of species pairs; positivity of $\cor_l(x_i,x_j)-\cor_u(x_i,x_j)$; and positivity of $\Atot$.
<!--***DAN: Do some search and standardize the use of lower v. left-tail and upper v. right-tail
association, and related quantities
***Shya: I did, it is always lower-tail and upper-tail association. Left- and right- tail for skewness.-->
<!--***DAN: Having defined this new piece of terminology, make sure to use it.-->

<!--omitted as of now: Significance tests for individual and overall asymmetry, tailsignif.R, binomial_sigtest.R
At present not shown, because we want to consider all interactions: weekest to strongest.

Tail-asymmetry could be weaker on individual species-pair, but on an aggregate level can make 
the community strongly tail-asymmetric. To test significantly stronger 
association in one of its extreme tail for a given species-pair, we simulate $10000$ made-up 
species pair with same Spearman correlation but removing their tail-asymmetry. If 
estimated tail-asymmetry $\cor_{l}-\cor_{u}$ on given data lies outside 
the $95\%$ confidence interval of the ones estimated with $10000$ surrogated data, then 
we consider the tail-asymmetric relationship is significant for that species-pair. 
On a community level, we calculated the fraction of significant species-pair among all correlated 
pairs considered (overall significance on the community level, $C_s$). We also perform a binomial 
test to check whether the community shows significantly stronger lower-tail (or upper-tail) association.
-->

<!--2. For Q2:
      - brief intro about the usual way to measure community variability (variance ratio method)
      - then introduce skewness ratio in analogy with variance ratio
      - separate the contributions from tail association and usual correlation (Pearson) 
      - need schematic diagram to show the relation (this diagram will again help to interpret results for Q2 later)
      - somewhere, in results or in discussion we should talk about the aggregated effect of weak interactions
-->

To accomplish goal G2 of the Introduction, we need a statistical measure of the distribution $\xtot$ that
characterizes community variability in a way that complements the classic use of $\var(\xtot)$ or $\CVcomsq$. We use
the skewness, denoted $\sk(\xtot)$. The skewness of $\xtot$ is the third central moment, 
$m_3(\xtot)$, divided by the cube of the standard 
deviation of $\xtot$. The
unbiased estimator (with Bessel's correction) was used for the standard deviation. For the third central moment, we used 
the third $h$-statistic, $m_3(\xtot)=\frac{T}{(T-2)(T-1)}\sum_{t=1}^T (\xtot(t)-\mean(\xtot))^3$, which is the 
unique symmetric unbiased estimator of that quantity. As illustrated in Fig. \ref{fig_pedag_cv2_skw}, skewness 
complements the traditional use of 
$\var(\xtot)$ or $\CVcomsq$ to characterize community variability, because strongly negative values indicate that $\xtot$
undergoes occasional large downward departures from typical values (i.e., crashes), and large positive values
indicate that $\xtot$ undergoes occasional large upward departures from typical values. As also illustrated in
Fig. \ref{fig_pedag_cv2_skw}, either of these dynamical behaviors can happen independently of the 
values of $\var(\xtot)$ and $\CVcomsq$. 

Just as, in the classic approach, the variance ratio $\phi_{\text{cv}}$ is the quotient $\CVcomsq/\CVindsq$, we 
analogously define a *skewness ratio*, as follows. Let $\Scom=\sk(\xtot)$, and consider the value that 
this quantity would take if species dynamics were independent.
We have $\sk(\xtot)=\frac{m_3(\xtot)}{\var(\xtot)^{3/2}}=\frac{\sum_{i,j,k}m_{ijk}}{(\sum_{i,j}v_{ij})^{3/2}}$,
where $m_{ijk}=\frac{T}{(T-2)(T-1)}\sum_{t=1}^T (x_i(t)-\mean(x_i))(x_j(t)-\mean(x_j))(x_k(t)-\mean(x_k))$ is a third
cross-moment. Assuming independence of species dynamics, this becomes $\Sind=\frac{\sum_i m_{iii}}{(\sum_i v_{ii})^{3/2}}$.
The *skewness ratio* is $\phi_S=\Scom/\Sind$, so that $\Scom=\phi_S \Sind$. 

Values of the skewness ratio provide a summary of how species interactions influence community variability that 
complements the information provided by the variance ratio. The skewness ratio can be positive or negative because 
$\Scom$ and $\Sind$ can be positive or negative, unlike $\CVcomsq$ and $\CVindsq$. If $\phi_S>1$, 
$\Scom$ has greater magnitude, but the same sign, as what its value would be in 
the absence of species interactions, i.e., $\Sind$. In such a case, and assuming for simplicity $\phi_{\text{cv}} \approx 1$
(see Discussion for more on this assumption), species interactions
magnify the tendency for the community aggregate quantity $\xtot$ to crash, or explode to large values.
This can be viewed as a new type of synchrony among species, in the sense that species relationships
inflate community instability by accentuating either positive or negative skewness. It is the
mutual sign of $\Scom$ and $\Sind$ that determines whether crashes or 
explosions to high values are magnified. 
If $0<\phi_S<1$, then $\Scom$ has lower magnitude, but the same sign, as $\Sind$, so species interactions
mitigate the tendency for the $\xtot$ to crash, or explode to large values (again assuming, for simplicity, that
$\phi_{\text{cv}} \approx 1$ -- see Discussion). This can be viewed as a new type of compensatory dynamics, in the sense that species 
relationships reduce community instability. 
<!--If $\phi_S \approx 0$,
then skewness of $\xtot$ is eliminated by species interactions, i.e., the tendency of the community to
asymmetrically crash or explode to high values is eliminated by species interactions.--> 
If $\phi_S<0$, then whereas, without species interactions, $\xtot$ would have exhibited occasional
crashes (or, respectively, explosions to high values), with species interactions $\xtot$ instead
exhibits occasional explosions to high values (or, respectively, crashes). 
<!--***DAN: Don't forget to add the promised Discussion material mentioned in this paragraph-->

Our skewness ratio does not supplant the classic variance ratio approach, but rather
complements it. Just as one understands more about a distribution by knowing both its variance and skewness, so
one understands more about community dynamics by combining variance ratio and skewness
ratio approaches. The values of $\phi_{\text{cv}}$ and $\phi_S$ provide complementary
information about how the distribution of $\xtot$ is influenced by species interactions (see Discussion).
See Fig. \ref{fig_method_box}A,B for a summary of the two approaches, and 
Fig. \ref{fig_method_box}E for a summary of notation.
<!--***DAN: Don't forget to add the promised Discussion material mentioned in this paragraph-->

<!--how phi_s and phi_cv are different?-->
<!--***DAN: Some of the below might be useful for the Discussion
We recommend to use both these indices ($\phi_{cv}$ and $\phi_s$) to get 
a complete picture to predict the stability of the community. It is possible
to underestimate the stability status of the community if we only
depend on the $\phi_{cv}$, which is a correlation 
(i.e. normalized covariance) based measure. $\phi_s$, on the other hand, takes 
the extreme values or tail asymmetry between interspecific 
response into account to assess the change in community structure
over time. Therefore, two communities with similar $\phi_{cv}$ could have two drastically 
different $\phi_s$ and interpretation of stability if we consider the change of skewness as an indicator 
of community-change (examples with real data are shown in \nameref{Results}). For example, consider 
two communities $C_A$ and $C_B$ with equal number of species. It is 
possible that the majority of species in $C_A$ show very low productivity for some 
years whereas those in $C_B$ are extremely abundant for some years, though
their overall pairwise correlation remains same. In such case, $C_A$ and $C_B$ 
should show more or less same $\phi_{cv}$ but 
different $\phi_s$, particularly $\phi_s$ for community $C_A$ 
should be less than that for community $C_B$, i.e., $0<\phi_{s,C_A}<1 < \phi_{s,C_B}$. 
***DAN: double check this statement, this is theoretically I expect but in Fig.1 with madeup extreme tail dep community phi_s are same though S_com are different
To explain the theoretical basis of such difference 
between $\phi_{cv}$ and $\phi_s$, we make partition for the overall contribution 
in stability metric into two distinct parts - due to correlation and
due to tail association for extreme values between interspecific response, as follows. 
-->

The skewness ratio $\phi_{S}$ provides information about the influence of species interactions, of all kinds,
on the skewness of $\xtot$; this is because $\phi_{S}$ is the quotient of $\Scom$, which is influenced by species interactions
of all kinds, and $\Sind$, which is influenced by none. But goal G3 of the Introduction was 
to understand the specific influence of species tail associations
and asymmetries of tail associations on $\Scom=\sk(\xtot)$. We did this through a randomization procedure
that rendered symmetric the tail associations between species, while keeping other statistical properties of 
community dynamics approximately fixed. Given the data $x_i(t)$ ($i=1,\ldots,N$, $t=1,\ldots,T$), the randomization 
procedure produced $M$ surrogate datasets $x_i^{(m)}(t)$ ($i=1,\ldots,N$, $t=1,\ldots,T$), for $m=1,\ldots,M$, 
for any desired number of surrogates, $M$, with properties detailed below. Since these surrogate datasets
had symmetric tail associations between species but were otherwise statistically similar to the original dataset,
$x_i(t)$, $\Scom=\sk(\xtot)$ was computed for the original data, and a surrogate skewness value, $\sk(\sum_i x_i^{(m)})$,
was also computed for each of the surrogate datasets. A test of whether asymmetry of tail associations significantly influenced
$\Scom$ was obtained by examining whether $\Scom$ fell sufficiently in the tails of the distribution 
of surrogate values to meet a desired statistical confidence level. 

The quantity $\Snta$ was defined as the median of the values $\sk(\sum_i x_i^{(m)})$,
and it is $\Snta$, and associated quantities which we will now define, that are used to accomplish goal G3 of the Introduction. 
We also defined the quantities $\phi_{S,\text{ta}}=\Scom/\Snta$ and $\phi_{S,\text{cor}}=\Snta/\Sind$,
where "ta" stands for "tail association" and "nta" stands for "no tail association".
The quantity $\phi_{S,\text{cor}}$ satisfies $\Snta=\phi_{S,\text{cor}}\Sind$, and therefore
represents the factor by which correlations between species, as distinct from asymmetries of tail association,
influence the skewness ratio. The subscript "cor" is a reference to the effects of correlation.
The quantity $\phi_{S,\text{ta}}$ satisfies $\Scom=\phi_{S,\text{ta}}\Snta$, and therefore
represents the factor by which asymmetric tail associations between species further influence community variation. 
So a value of $\phi_{S,\text{ta}}$ which differs from $1$ indicates that asymmetric tail associations have a role in
determining $\Scom$; statistical significance, here, is judged as in the previous paragraph.
It is straightforward to show from the definitions that $\phi_S=\phi_{S,\text{ta}}\phi_{S,\text{cor}}$, so
$\Scom=\phi_{S,\text{ta}}\phi_{S,\text{cor}}\Sind$.
Thus the two newly defined quantities, $\phi_{S,\text{ta}}$ and $\phi_{S,\text{cor}}$, separate the influence of 
species relationships on the skewness ratio into the factors due to asymmetries of tail associations and correlations
with symmetric tail associations. 
For comparison, $\CVcomsq$ was also computed for the empirical data and for surrogates, 
defining quantities $\CVntasq$, $\phi_{\text{cv},\text{cor}}$ and $\phi_{\text{cv},\text{ta}}$
which are related to each other in an analogous way. Fig. \ref{fig_method_box}A-D summarize the newly defined
quantities and their interrelationships. Because $\CVcomsq$ depends only 
on species means and covariances, which are preserved by surrogates, we should have $\CVntasq \approx \CVcomsq$ and 
$\phi_{\text{cv},\text{ta}} \approx 1$.

Having defined methods, we can now frame hypotheses suggested by Fig. \ref{fig_pedag_cv2_skw} and the 
Theory section in terms of the new methods; tests of the hypotheses will help address goal G3 of the Introduction. 
Fig. \ref{fig_pedag_cv2_skw}A,C,E and 
Fig. \ref{fig_theory}M-P suggest that communities exhibiting a preponderance of lower-tail association 
should have total community dynamics that are more left-skewed, or less right-skewed, than would have been the case
without these tail associations. Using the language of our new methods, this constitutes
a hypothesis that for communities for which $n_L>n_U$ and $\Atot>0$, we should have 
$\Scom<\Snta$. Fig. \ref{fig_pedag_cv2_skw}B,D,F and 
Fig. \ref{fig_theory}I-L suggest that communities exhibiting a preponderance of upper-tail association 
should have total community dynamics that are more right-skewed, or less left-skewed, than would have been the case
without these tail associations. Using the language of our new methods, this constitutes
a hypothesis that for communities for which $n_L<n_U$ and $\Atot<0$, we should have 
$\Scom>\Snta$.
<!--***DAN: Don't forget to make comparisons in the Results between the values of phi_S,ta and phi_S,cor,
to judge the relative importance of asymmtric tail associations and other aspects of species interrelationships.
Also, relate the results to stability, which is not conveniently done here because it depends on signs as
well as magnitudes, e.g., if Scom<Snta and both are negative, then tail association have destabilized the community, I think,
i.e., crashes to low values are accentuated or something. Better to work all these specifics out for the cases
that actually materialize than to try to consider all possibilities here-->

The surrogate datasets, $x_i^{(m)}(t)$, have the following properties. Complete descriptions of surrogates are in SI
section \ref{SI-surrogs}. 
First, for each $m$ and $i$, the sets $\{x_i(t) : t=1,\ldots,T\}$ and 
$\{x_i^{(m)}(t) : t=1,\ldots,T\}$ are identical, as unordered sets.
In particular this implies that the temporal means and variances of the time series
$x_i(t)$ and $x_i^{(m)}(t)$ are identical, for all $i$ and $m$. 
Second, the expected value of the sum $\sum_{i_2>i_1} \cov(x_{i_1}^{(m)}(t),x_{i_2}^{(m)}(t))$
is within $\pm 10\%$ of the sum $\sum_{i_2>i_1} \cov(x_{i_1}(t),x_{i_2}(t))$;
and for any $i_1$ and $i_2$, the expected value of
$\cor(x_{i_1}^{(m)}(t),x_{i_2}^{(m)}(t))$ is as close as possible to
$\cor(x_{i_1}(t),x_{i_2}(t))$ (see SI section \ref{SI-surrogs} for information on the extent to which
these quantities agreed for our datasets).
<!--***DAN: Once you put all the relevant material and figures in the sup mat, you may want to change this 
reference to it is a reference to the appropriate SI figure instead of or in addition to the SI section on surrogs-->
In particular, this means that the total community variance, 
$\var(\sum_i x_i^{(m)}(t))= \sum_{i_1, i_2} \cov(x_{i_1}^{(m)}(t),x_{i_2}^{(m)}(t))$
for a surrogate dataset should be similar to the empirical total community variance,
$\var(\sum_i x_i(t))= \sum_{i_1, i_2} \cov(x_{i_1}(t),x_{i_2}(t))$.
Likewise,
the variance ratios $\frac{\sum_{i_1, i_2} \cov(x_{i_1}^{(m)}(t),x_{i_2}^{(m)}(t))}
{\sum_{i} \var(x_i^{(m)}(t))}$ and 
$\frac{\sum_{i_1, i_2} \cov(x_{i_1}(t),x_{i_2}(t))}
{\sum_{i} \var(x_i(t))}$ should be similar.
Finally, for any $i_1$, $i_2$ and $m$, the tail associations of $x_{i_1}^{(m)}(t)$
and $x_{i_2}^{(m)}(t)$ are symmetric, i.e., similar in their upper- and lower-tail association
values. All analyses were done in R. Complete computer code for this project is at https://github.com/sghosh89/Copula_spaceavg. 


```{r pval_skw, echo=F}
hays_skwres<-readRDS("./Results/hays_results/skewness_results/stability_CP_hays.RDS")
pval_skw_hays<-sum(hays_skwres$skw_surrogs<hays_skwres$df_stability$skw_real)/length(hays_skwres$skw_surrogs)

knz_skwres<-readRDS("./Results/knz_results/skewness_results/stability_CP_knz.RDS")
pval_skw_knz<-sum(knz_skwres$skw_surrogs>knz_skwres$df_stability$skw_real)/length(knz_skwres$skw_surrogs)
```

# Results\label{Results}

<!--***DAN: Shyamolina, I have thought about how I am going to edit this Results section, and I've made a plan, but I need
a few things from you to implement that plan. 

Please compute the fraction of surrogate values on the current Fig. 6F (Dan, this is the new Fig. S5B) that are to the *left* of the black dot. This is the 
p-value for a one-tailed test. It looks like it should be a very small number. Please type this number in a reponse to this comment, 
but *also* please make the number available as an R variable, so when I edit the results I can render the number using an autolink. 
Please say what the R variable name is in a response to this comment. 
***Shya: 0.0027, variable name you can use: pval_skw_hays 

Please compute the fraction of surrogate values on the current Fig. 6H that are to the *right* of the black dot (Dan, this is the current Fig. S5D). This is the 
p-value for a one-tailed test. It looks like it should be a small number. Please type this number in a reponse to this comment, 
but *also* please make the number available as an R variable, so when I edit the results I can render the number using an autolink. 
Please say what the R variable name is in a response to this comment. 

***Shya: 0.0252, variable name you can use: pval_skw_knz 
-->

Addressing goal G1 of the Introduction, the quantities $\cor_l-\cor_u$ for all positively associated 
pairs of species within the Hays and Konza datasets are displayed in Fig. \ref{fig_CorlmCoru}, as are the
community aggregate quantities $n_L$ and $n_U$ and the total community tail association values $\Atot$ (Methods). 
Although both the Hays and Konza datasets had pairs of species with more 
lower-tail association and pairs with more upper-tail association, $n_L$ was greater than $n_U$ and $\Atot$ was positive
for Hays, and $n_L$ was less than $n_U$ and $\Atot$ was negative for Konza (Fig. \ref{fig_CorlmCoru}). Thus 
lower-tail association was dominant at Hays and upper-tail association was dominant at Konza.

We earlier framed the hypotheses that communities exhibiting more lower-tail association 
should also have $\xtot$ that is more left-skewed, or less right-skewed, than would have been the case
without species tail associations; and that communities exhibiting more upper-tail association 
should have $\xtot$ that is more right-skewed, or less left-skewed, than would have been the case
without tail associations (Methods). We tested these hypotheses, thereby addressing 
goals G2 and G3 of the Introduction. The quantities of Fig. \ref{fig_method_box} are displayed for our 
two datasets in Fig. \ref{fig_result_box}.
Because $n_L>n_U$ and $\Atot>0$ for Hays (Fig. \ref{fig_CorlmCoru}A), $\Scom$ should be less than $\Snta$.
This was confirmed (Fig. \ref{fig_result_box}B). Both $\Scom$ and $\Snta$ were positive and the component 
of the skewness ratio due to tail association, $\phi_{S,\text{ta}}$, was less than 1,
i.e., $\phi_{S,\text{ta}}=0.2757<1$. 
<!--***DAN: Shyamolina, please make the above number an autolink-->
Becuse $n_L<n_U$ and $\Atot<0$ for Konza (Fig. \ref{fig_CorlmCoru}B), $\Scom$ should be greater than 
$\Snta$. This was also confirmed (Fig. \ref{fig_result_box}D). $\Scom$ and $\Snta$ were again positive,
so for Konza $\phi_{S,\text{ta}}$ was greater than 1 (it was $2.7999$). 
<!--***DAN: Shyamolina, please make the above number an autolink-->
These results were statistically significant: for Hays, $\Scom$ was less than a fraction $`r round(1-pval_skw_hays,4)`$
of the quantities $\sk(\sum_i x_i^{(m)})$ for the Hays surrogates $m=1,\ldots,M$ (Methods);
whereas for Konza, $\Scom$ was greater than a fraction $`r round(1-pval_skw_knz,3)`$ of the analogous surrogate 
values for Konza (Fig. \ref{SI-fig_result_box}). Thus asymmetric 
tail associations between species can 
significantly modify the skewness of the total community abundance, $\xtot$, either raising it (upper-tail asociations)
it or lowering it (lower-tail associations).  

As anticipated (Methods), $\CVntasq \approx \CVcomsq$ and $\phi_{\text{cv,ta}} \approx 1$ for both Hays and Konza 
(Fig. \ref{fig_result_box}A, C) -- the variance ratio approach does register the importance of 
tail associations for community variability. This was expected because $\CVcomsq$ 
depends only on species means and covariances, which were preserved 
by surrogates (Methods). The distribution, across surrogates $m=1,\ldots,M$, of the squared coefficients of variation of the quantities 
$\sum_i x_i^{(m)}$ was centered very close to the value of $\CVcomsq$, for both Hays and Konza (Fig. \ref{SI-fig_result_box}A, C).

We compare conclusions about synchronous versus compensatory dynamics that would typically be drawn about Hays using
only the variance ratio approach to conclusions drawn using both approaches together.
Because $\phi_{\text{cv}}<1$ (Fig. \ref{fig_result_box}), the conclusion solely based on the classic 
variance ratio approach would simply have been 
that population dynamics at Hays are compensatory, i.e., species interactions reduce or buffer against
community instability. However, the combined approach yields more nuanced information.  
The skewness ratio, $\phi_S$, was substantially less than $1$ 
(it was $0.2911$), with $\Scom=0.2246$ less than $\Sind=0.7718$ (Fig. \ref{fig_result_box}).  
<!--***DAN: Shyamolina, please make the above three numbers into autolinks-->
So the tendency for the total community abundance $\xtot$ to explode to high values (right skewness)
is mitigated by species interactions. This new information agrees with and supports the old, variance-ratio-based
conclusion that species interactions are generally stabilizing in the Hays community, but the new information
also complements and extends the classic conclusion by revealing 
the greater importance of community interactions for mitigating explosions of $\xtot$ to high values,
compared to a lesser role in mitigating crashes of $\xtot$ to low values. Apparently the role of species interactions
in mitigatingor accentuating extreme values of $\xtot$ can differ between extremely low versus high values of $\xtot$. 
See SI section \ref{ecdf_approach} for more results
and further discussion on the relative importance of community interactions for mitigating community
crashes versus explosions in Hays.

We next compare conclusions about synchronous versus compensatory dynamics that would typically be drawn about Konza using
only the variance ratio approach to conclusions drawn using both approaches.
Because $\phi_{\text{cv}}$ was slightly less than $1$ for Konza (Fig. \ref{fig_result_box}), the conclusion solely based on the classic 
variance ratio approach would have been 
that population dynamics at Konza are slightly compensatory, i.e., species interactions slightly reduce or buffer against
community instability. But using both the variance and skewness ratios together yields conclusions that differ 
substantially. The skewness ratio, $\phi_S$, for Konza was much greater than $1$ 
(it was $7.3811$), with $\Scom=1.3655$ correspondingly much greater than $\Sind=0.1850$ (Fig. \ref{fig_result_box}).  
<!--***DAN: Shyamolina, please make the above three numbers into autolinks-->
Thus, although species interactions slightly reduce the variance of $\xtot$, they also dramatically increase right skew
of $\xtot$, leading to an overall *greater* propensity for explosions of $\xtot$ to high values. Crashes of $\xtot$
to low values are reduced by species interactions. The classic, variance ratio result obscures the fact that species interactions
modify the extreme bahavior of $\xtot$ very differently in the two extremes of its distribution. See SI section \ref{ecdf_approach} 
for more results and further discussion.

Combining the skewness and variance ratio approaches also reveals differences between Hays and Konza that the variance 
ratio alone does not  reveal. For both communities, $\phi_{\text{cv}}$ is slightly or moderately less than $1$,
suggesting compensatory, stabilizing species interactions in both communities. But $\phi_S$ was less than 1 for Hays 
and was markedly greater than 1 for Konza. The influence of tail associations between species on the distribution 
and extreme values of $\xtot$, not revealed by the classic variance ratio approach, differs markedly between the Hays 
and Konza systems. 

Our results also reveal that the influence of tail associations between species on the distribution of $\xtot$ is
just as great as the influence of aspects of species interactions other than tail association. For Konza, 
$\phi_{S,\text{cor}}$ and $\phi_{S,\text{ta}}$ were similar (Fig. \ref{fig_result_box}), indicating that both tail associations and
aspects of species interactions that were not related to tail associations were comparable in their effects on the skewness
of $\xtot$ (recall that $\Scom=\phi_{S,\text{ta}} \phi_{S,\text{cor}} \Sind$). However, for Hays, $\phi_{S,\text{cor}}$
was close to 1 and $\phi_{S,\text{ta}}$ differed substantially from 1 (Fig. \ref{fig_result_box}), indicating that essentially all of the 
influence of species interactions on the skewness of $\xtot$ is due to tail associations. 

<!--***DAN: Your capitalization of CV is inconsistent across the vaious notation (see \phi_cv and related quantities, compared to
CV_com^2 and related quantities) and in Fig. 6. Likewise for S. Fix it by just capitalizing CV everywhere and S everywhere. You also
need S to be italicized everywhere, and CV to be upright everywhere.-->

# Discussion\label{Discussion}




# Acknowledgments
\noindent The authors thank Jim Bever, Bryan Foster and Lauren Hallett for helpful discussions. The authors were partly supported 
by U.S. National Science Foundation grant 1714195, the James S. McDonnell Foundation, and the California Department
of Fish and Wildlife Delta Science Program. 

<!--Pedagogical figure-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=18cm]{./Results/pedagog_figs/pedagog_cv2_skw.pdf}
\caption{Pedagogical example showing two communities (A, B) of twenty species each that are not distinguishable 
via the classic variance ratio approach and related approaches, but that nonetheless differ markedly in 
potentially important ways. Species marginal distributions, and therefore species means and variances
over time, were the same in both scenarios, as were species covariances, up to sampling variance. 
Thus the quantities $\CVcomsq$, $\CVindsq$ and $\phi_{\text{cv}}$ of the classic variance ratio approach
were essentially the same in the two scenarios (A, B), as was the Loreau-de Mazancourt (Loreau \emph{et al.} 2008)
variance ratio, denoted $\phi_{\text{LdM}}$. Nevertheless, differences between the scenarios in interspecific relationships, not 
captured by covariances, resulted in differently skewed total time series, $\xtot(t)$, in the two scenarios
(C-F). The total was more likely to crash to low values in the first scenario (C, E)
than in the second scenario (D, F; compare the probabilities on E and F listed to the left of the dashed line), 
whereas $\xtot(t)$ was more likely to explode to high values 
in the second scenario than in the first (probabilities on E, F to the right of the dotted line). 
Dashed and dotted lines show lower and upper thresholds which may be considered
disasterous for $\xtot(t)$ to cross. Probabilities of crossing these thresholds were negligible for one scenario, but
non-negligible for the other.
In both scenarios, there were two groups of species (red and black lines).
Species in different groups were compensatory, but for scenario 1, within-group synchrony occurred principally
when populations were low, and for scenario 2 when populations were high. See the Introduction and SI section \ref{SI-PedagogFigMethods} 
for notation and 
additional details. Time series were generated and statistics computed for 100000 years, but are plotted on A, B
for 60 years only, for clarity. \label{fig_pedag_cv2_skw}}
\end{center}
\end{figure}

<!--Theory figure-->
\begin{figure}
\begin{center}
\includegraphics[width=9cm]{./Results/TheoryFig.pdf}
\caption{Four simple theoretical scenarios demonstrating how species interrelationships can influence a community aggregate
quantity in ways that are not detectable to classic approaches. See the Theory section for interpretations and SI section 
\ref{SI-TheoryMethods} for mathematical details.\label{fig_theory}}
\end{center}
\end{figure}

<!--Methods figure for asymmetric tail association and partial Spearman correlation-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/pedagog_figs/LTUT_rho_0.8.pdf}
\caption{Normalized rank plots visually reveal that a positive association is asymmetric in the tails if 
points are clustered closer to the $u_i=u_j$ diagonal in the upper right or lower left corners. 
The partial Spearman correlation, $\cor_{l_b,u_b}(x_i,x_j)$, within a band defined by $l_b$ and $u_b$
(see Methods) can be computed for any band to quantify the component of the overall Spearman correlation
due to that band. Solid diagonal lines show two alternative bands. Lower-tail association is greater than upper-tail 
association on (A), so the partial Spearman correlation using the band circumscribing the
red points will be greater on (A) than for the band around the blue points. Reverse statements hold on (B).
The (total) Spearman correlation is the same on both panels, $0.8$. 
This figure was slightly modified from Fig. 4 of 
Ghosh \emph{et al.} 2020a and Fig. 7 of Ghosh \emph{et al.} 2020b. \label{fig_pedag_taildep}}
\end{center}
\end{figure}
<!--***DAN: This is a reminder that the citations in the above caption are not autolinks, so need to be
changed manually if anything changes. They are supposed to be to BIVAN and the extinction risk paper. Do
not delete this reminder.-->
<!--***DAN: Here is some stuff that I cut from the above figure, keeping in case needed later:
\includegraphics[width=13cm]{./Results/pedagog_figs/pedagog_simple_cv2_skw.pdf}\\
Upper panel shows a schematic diagram of the known $(A)$ variance-ratio approach usually used
to assess community-stability and our proposed $(B)$ skewness-based analog; \textit{CV} and \textit{S} denote 
coefficient of variation and skewness, respectively. White boxes with dashed arrows indicate a missing link between 
the two extremes: an actual community ($com$) having both correlation and tail 
association between species and its independent ($ind$) analog with no
interspecific interaction. This is explored in detail later (Fig. \ref{fig_method_box}).-->

<!--Pedagogical figure to explain method-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.5in]{./Results/pedagog_figs/method_fig.pdf}\\
\caption{Summary of approach and notation. Arrows in A-D imply an equation
where the quantity in the box at the arrow tail, times the quantity labelled on 
the arrow itself, equals the quantity in the box at the arrow head. For instance, the 
arrow in A represents the quation $\CVcomsq=\phi_{\text{cv}} \CVindsq$. \label{fig_method_box}}
\end{center}
\end{figure}

<!--Results with corl-coru matrix plot-->
\begin{figure*}[!h]
\begin{center}
\textbf{ \hspace{-7 cm} (A) \hspace{7 cm} (B)}\\
\includegraphics[width=8cm]{./Results/hays_results/corstat_hays_spaceavg_results/Corl-Coru_ub_0.5.pdf}
\includegraphics[width=8cm]{./Results/knz_results/corstat_knz_spaceavg_results/soiltype_t/Corl-Coru_ub_0.5.pdf} 
\caption{Measures of asymmetry of tail association for all pairs of species with positive Spearman correlation,
using the Hays (A) and Konza (B) datasets. The quantity $\cor_l-\cor_u$ (Methods) is displayed, using a blue/red split colorbar,
for all positively associated species. Green-shaded species pairs were negatively associated. See
Tables \ref{SI-tab_spinfo_hays}-\ref{SI-tab_spinfo_knz} for species names. Although both lower- and upper-tail
association occur in both datasets, the quantities $n_L$, $n_U$ and $\Atot$ (Methods), displayed under each 
panel, indicate that lower-tail association between species was stronger or more common in the Hays data, and 
upper-tail association was stronger or more common in the Konza data. \label{fig_CorlmCoru}}
\end{center}
\end{figure*}

<!--Skewness results for both datasets-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=14cm]{./Results/pedagog_figs/Result_box.pdf}
\caption{The quantities of Fig. \ref{fig_method_box} computed for the Hays (A, B) and Konza (C, D)
datasets. See SI Fig. \ref{SI-fig_result_box} for related results. \label{fig_result_box}}
\end{center}
\end{figure}

\clearpage

# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{1pt}
\noindent








